---
title: "Reproducible Research PA1.Rmd"
author: "Xiuyi He"
date: "August 8, 2015"
output: html_document
---

Notes:First of all, because my html file is too big to shown in the github website,
I write some code at the botton in this Rmd to save the pictures.
I will also upload these pictures for viewing and grading.
Sorry for the inconvenience and thanks a lot!

**Loading and preprocessing the data**

```{r echo=TRUE}
data <- read.csv("activity.csv")
head(data)
str(data)
```


**What is mean total number of steps taken per day?**
1.Calculate the total number of steps taken per day
```{r echo=TRUE}
tapply(data$steps,data$date,sum)
```

2.Make a histogram of the total number of steps taken each day
```{r echo=TRUE}
stepsum <- tapply(data$steps,data$date,sum)
hist(stepsum)
```

3.Calculate and report the mean and median of the total number of steps taken per day
```{r echo=TRUE}
mean(stepsum[!is.na(stepsum)])
median(stepsum[!is.na(stepsum)])
```

**What is the average daily activity pattern?**

1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo=TRUE}
sub <- data[complete.cases(data$steps),]
data$interval <- as.numeric(data$interval)
for(i in 1:17568){
      if(is.na(data$steps[i])){
             data$steps[i] <- mean(sub$steps[sub$interval == data$interval[i]])
       }
 }
meansteps <- tapply(data$steps,data$interval,mean)
data <- read.csv("activity.csv")
interval <- data$interval[1:288]
combine <- data.frame(cbind(interval,meansteps))
plot(combine$interval,combine$meansteps,type="l",xlab="the 5-minute interval",ylab="the average number of steps taken, averaged across all days",main="the time series plot")
```

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
maxvalue <- max(meansteps)
subcombine <- combine[combine$meansteps==maxvalue,]
subcombine$interval
```


**Imputing missing values**
1.Calculate and report the total number of missing values in the dataset
```{r echo=TRUE}
length(data$steps[is.na(data$steps)])
```

2.Devise a strategy for filling in all of the missing values in the dataset.
I will use the mean of that interval to fill out the NAs
```{r echo=TRUE}
sub <- data[complete.cases(data$steps),]
```
3.then write a function to fill out the NAs
```{r echo=TRUE}
data$interval <- as.numeric(data$interval)
for(i in 1:17568){
      if(is.na(data$steps[i])){
            data$steps[i] <- mean(sub$steps[sub$interval == data$interval[i]])
      }
}
head(data)
```
See? Here is the new dataset!

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r echo=TRUE}
newstepsum <- tapply(data$steps,data$date,sum)
hist(newstepsum)
mean(newstepsum)
median(newstepsum)
```

**Are there differences in activity patterns between weekdays and weekends?**
```{r echo=TRUE}
library(dplyr)
library(lubridate)
data1 <- mutate(data,weekday = wday(ymd(data[,2])))
data1$weekday <- as.numeric(data1$weekday)
for(i in 1:17568){
      if(data1[i,4] == 1){
            data1[i,4] <- "weekend"
      }else if (data1[i,4] == 7){
            data1[i,4] <- "weekend"
      }else {
            data1[i,4] <- "weekday"
      }
}
data1$weekday <- as.factor(data1$weekday)
```

the plot

```{r echo=TRUE}

#process the data set
subweekday <- data1[data1$weekday == "weekday",]
subweekend <- data1[data1$weekday == "weekend",]
pweekday <- aggregate(subweekday$steps,by=list(subweekday$interval),FUN=mean)
colnames(pweekday) <- c("interval","meansteps")
pweekday$fweekday <- "weekday"
pweekend <- aggregate(subweekend$steps,by=list(subweekend$interval),FUN=mean)
colnames(pweekend) <- c("interval","meansteps")
pweekend$fweekday <- "weekend"
data2 <- rbind(pweekday,pweekend)


#plot the picture
library(lattice)
xyplot(meansteps~interval|fweekday,data2,type="l",layout=c(1,2))
```





